---
title: "Module 1"
author: "Yu Zhang"
date: "2021/11/15"
output: html_document
---

1.1 R can be used as a calculator

```{r}
1+1
5*(4-1)^2
sqrt( log(10) )
```

1.2 Installing packages in R

```{r}
# install.packages('ggplot2')
```

```{r}
library('ggplot2')
```

```{r}
getwd()
```
```{r setup, include=FALSE, echo=FALSE}
require("knitr")
# basedirectory = 'C://Users//yzhan//Dropbox//Teaching教学//金融硕金融计量//'
basedirectory = 'D://Dropbox//Teaching教学//金融硕金融计量//'
subdirectory = 'Rfiles//'
workdirectory = paste(basedirectory, subdirectory, sep = '')
opts_knit$set(root.dir = workdirectory)
```

```{r}
getwd()
```

1.3 Input your first column into R

```{r}
 (a <- c(1,2,3,4,5,6))
```

```{r}
(year <- c(2016,2017,2018,2019,2020))
(stkprc1 <- c(10,20,40,80,160))
(stkprc2 <- c(40,30,20,10,5))
(portfolio <- 1*stkprc1 + 1*stkprc2)
(price_difference <- stkprc1 - stkprc2)
length(year)
sort(stkprc2)
rep(1,5)
seq(4,12,2)
(cities <- c("New York","Los Angeles","Chicago"))
(cities == "New York")
```

Factor variables

```{r}
x <- c(3,2,2,3,1,2,3,2,1,2)
(xf <- factor(x, labels=c("C","B","A")))
```
```{r}
credit_rating <- c("AAA", "AA", "A", "BBB", "AA", "BBB", "A")
(credit_factor <- factor(c("AAA", "AA", "A", "BBB", "AA", "BBB", "A")))
(credit_factor_ordered <- factor(credit_rating, ordered = TRUE, 
                                levels = rev(c("AAA", "AA", "A", "BBB")))) # rev() is for reverse
(credit_factor_ordered2 <- ordered(credit_factor, levels = rev(c("AAA", "AA", "A", "BBB")))) # rev() is for reverse
```

Vector naming

```{r}
# Create a vector "avgs":
(avgs <- c(.366, .358, .356, .349, .346))
```

```{r}
# Create a string vector of names:
(players <- c("丁一","钱二","张三","李四","王五"))
```

```{r}
# Assign names to vector and display vector:
names(avgs) <- players
avgs
names(avgs)
```
```{r}
# Indices by number:
avgs[2]
avgs[1:4]
avgs[-(1:3)]
avgs[4:length(avgs)]
```
```{r}
# Indices by name:
avgs["王五"]
```
```{r}
players=="王五"
avgs[players=="王五"]
stkprc1[year>=2018]
stkprc1_named <- stkprc1
names(stkprc1_named) <- year
stkprc1_named[year>=2018]
```

```{r}
# Logical indices:
avgs[ avgs>=0.35 ]
```

R vector to R matrix

```{r}
(v <- c(2,-4,-1,5,7,0))
( A <- matrix(v,nrow=2) )
( matrix(v,ncol=2) )
( matrix(v,ncol=4) )
(A)
(A[1,])
(A[,1])
```

```{r}
(B <- cbind(year, stkprc1, stkprc2))
```
```{r}
B[1,]
B[year>=2017,]
B[,colnames(B)=='year']
B[,'year']
B[,colnames(B)==c('year','stkprc1')]
B[,c('year','stkprc1')]
```

R matrix to R dataframe

```{r}
(C <- as.data.frame(B))
```

```{r}
C[1,]
C[year==2017,]
C[,colnames(C)=='year']
C[,'year']
C$year
C[year>=2018,]
C[,colnames(C)==c('year','stkprc1')]
C[,c('year','stkprc1')]
C[year>=2019,colnames(C)==c('year','stkprc1')]
C[year>=2019,names(C)==c('year','stkprc1')]
C[year>=2019,!(names(C) == c('stkprc1'))]
C[year>=2019,!(names(C) %in% c('stkprc1'))]
```

```{r}
getwd()
save(C, file = "1_stkprc.RData")
load('1_stkprc.RData')
```

```{r}
write.table(C, file = "1_stkprc.csv", sep=",", row.names = FALSE)
```

```{r}
(D <- read.table("1_stkprc.csv", sep=",", header = TRUE))
```
```{r}
D
```

在R语言中调用Akshare的数据接口

首先官网安装python并在Powershell（Terminal）中执行pip install akshare

然后在R语言中安装reticulate

```{r}
if(!require(reticulate)){
    install.packages("reticulate")
}
py_config()
```

```{r}
library(reticulate)  # 导入 reticulate 包
ak <- import("akshare")  # 类似于 import akshare as ak
stock_snapshot <- ak$stock_zh_a_spot_em()  # 类似于 ak.stock_zh_a_spot_em()
print(stock_snapshot)  # 查看数据
```
```{r}
stock_history_qfq <- ak$stock_zh_a_hist(symbol='600519',period='daily',adjust='qfq')  # 前复权（当前价格不变）
stock_history_hfq <- ak$stock_zh_a_hist(symbol='600519',period='daily',adjust='hfq')  # 后复权（历史价格不变）
print(stock_history_qfq)
print(stock_history_qfq[order(stock_history_qfq$日期, decreasing = TRUE),])
print(tail(stock_history_qfq))
print(stock_history_hfq)
print(stock_history_hfq[order(stock_history_hfq$日期, decreasing = TRUE),c('日期','收盘')])
print(tail(stock_history_hfq))
```
```{r}
stock_history_hfq$名称 <- '贵州茅台'
stock_history_hfq <- stock_history_hfq[,c(ncol(stock_history_hfq),1:(ncol(stock_history_hfq)-1))]
stock_history_hfq$代码 <- '600519'
stock_history_hfq <- stock_history_hfq[,c(ncol(stock_history_hfq),1:(ncol(stock_history_hfq)-1))]
print(head(stock_history_hfq))
```

计算回报率

```{r}
stock_history_hfq$log_close_price <- log(stock_history_hfq$收盘)
stock_history_hfq$ret_i <- c(NA,diff(stock_history_hfq$log_close_price))
stock_history_hfq <- stock_history_hfq[,!(names(stock_history_hfq) %in% 'log_close_price')]
print(tail(stock_history_hfq))
stock_return <- stock_history_hfq[,c('名称','代码','日期','ret_i')]
print(stock_return)
```
Use Stargazer to summarize the stock return dataset

```{r}
if (!require('stargazer')) {
  install.packages('stargazer')
}
library('stargazer')
stargazer(stock_history_hfq, type="text")
stargazer(stock_return, type="text")
```
单数日和双数日回报是否相同？

```{r}
单数日 = strtoi(substring(stock_return$日期,10,10)) %% 2
(stock_return <- cbind(stock_return, 单数日))
```

```{r}
t.test(stock_return$ret_i ~ stock_return$单数日,paired=FALSE,var.equal=FALSE,alternative='two.sided') 
 # we use unpaired t test (paired t test is for same individual over time)
```

Plot the histogram and kernel density of ret_i

```{r}
library('ggplot2')
ggplot(stock_return, aes(x=ret_i))+geom_histogram(aes(y=..density..))+geom_density()
```

Plot the age distribution for males and females

```{r}
ggplot(stock_return, aes(x=as.factor(单数日), y=ret_i, fill=as.factor(单数日))) + geom_boxplot()
```

```{r}
library('dplyr')
(D <- C[,c('year','stkprc1')])
(E <- C[,c('year','stkprc2')])
(F <- left_join(D,E,by='year'))
```
```{r}
D
```

```{r}
E
```

```{r}
F
```

```{r}
date <- as.Date(stock_return$日期, format = "%Y-%m-%d")
```
