---
title: "Remainder of URfIE Chapter 1"
author: "FinEmt 2021"
date: "2021/11/21"
output: html_document
---

## 1.4 Base Graphics

In R, a function plot can be generated using the command **curve( function(x), xmin, xmax )**

```{r}
curve( x^2, -2, 2)
```
```{r}
curve( dnorm(x, mean=0.2, sd=0.4), -1, 1.5)
```

Base R plotting can also handle (1) overlays of multiple plots and (2) adding graph legends

```{r}
curve( dnorm(x,0,1), -10, 10, lwd=1, lty=1)
curve( dnorm(x,0,2),add=TRUE, lwd=2, lty=2)
curve( dnorm(x,0,3),add=TRUE, lwd=3, lty=3)
# Add the legend
# legend("topright",c("sigma=1","sigma=2","sigma=3"), lwd=1:3, lty=1:3)
# Add the legend with greek sigma
legend("topleft",expression(sigma==1,sigma==2,sigma==3),lwd=1:3,lty=1:3)
# Add the text with the formula, centered at x=6 and y=0.3
text(6,.3,
expression(f(x)==frac(1,sqrt(2*pi)*sigma)*e^{-frac(x^2,2*sigma^2)}))
```

## 1.5 "Tidyverse": `ggplot2` + `dplyr`

Tidyverse is a set of packages (e.g. `ggplot2` and `dplyr`) that makes the R workflow "pretty". 

### 1.5.1 `ggplot2`

First, let us use the `mpg` data set that comes with the package `ggplot2` to try out the command `ggplot`.

```{r echo=FALSE}
library(ggplot2)
```
```{r}
head(mpg)
```
A simple scatter plot (`geom_point`)
```{r}
ggplot() + geom_point( data=mpg, mapping=aes(x=displ, y=hwy) )
# ggplot(data=mpg, aes(x=displ, y=hwy)) + geom_point()
```
A scatter plot with regression line or smoothed line (`geom_point` + `geom_smooth`)
```{r}
# ggplot() + geom_point( data=mpg, mapping=aes(x=displ, y=hwy) ) + geom_smooth(data=mpg, mapping=aes(x=displ, y=hwy) )
ggplot(data=mpg, aes(x=displ, y=hwy)) + geom_point() + geom_smooth()
```
```{r}
# ggplot() + geom_point( data=mpg, mapping=aes(x=displ, y=hwy) ) + geom_smooth(data=mpg, mapping=aes(x=displ, y=hwy), method='lm', formula = y~x )
ggplot(data=mpg, aes(x=displ, y=hwy)) + geom_point() + geom_smooth(method='lm')
```

You can use `labs()` to modify the axis labels. 

```{r}
# ggplot() + geom_point( data=mpg, mapping=aes(x=displ, y=hwy) ) + geom_smooth(data=mpg, mapping=aes(x=displ, y=hwy), method='lm', formula = y~x )
ggplot(data=mpg, aes(x=displ, y=hwy)) + geom_point() + geom_smooth(method='lm') + labs(x="Engine displacement (1000cc)",y="Hwy mileage per gallon")
```

You can group the data using `color` or shape:

```{r}
ggplot(mpg, aes(x=displ, y=hwy)) + geom_point(aes(color=class))
```
```{r warning = FALSE, message=FALSE}
ggplot(mpg, aes(x=displ, y=hwy)) + geom_point(aes(shape=class))
```

```{r warning = FALSE, message=FALSE}
ggplot(mpg, aes(x=displ, y=hwy, color=class, shape=class)) + geom_point() + geom_smooth(se=FALSE) 
```

### 1.5.2 `dplyr`

Let us use the life expectancy (`SP.DYN.LE00.FE.IN`) data from the World Bank's World Development Indices to illustrate the `dplyr` package

```{r echo=FALSE}
library(WDI)
```

```{r eval=FALSE}
wdi_raw <- WDI(indicator=c("SP.DYN.LE00.FE.IN","SP.DYN.LE00.MA.IN"), start = 1960, end = 2014)
head(wdi_raw)
save(wdi_raw, file="wdi_raw.Rdata")
```


```{r}
load(file = "wdi_raw.Rdata")
```

```{r echo=FALSE}
library(dplyr)
```

```{r}
# filter: only China data
ourdata <- filter(wdi_raw, iso2c=="CN")
# rename lifeexpectancy variables
ourdata <- rename(ourdata, LE_fem=SP.DYN.LE00.FE.IN)
ourdata <- rename(ourdata, LE_male=SP.DYN.LE00.MA.IN)
# select relevant variables # help(select, package='dplyr')
ourdata <- select(ourdata, year, LE_fem, LE_male)
# order by year (increasing)
ourdata <- arrange(ourdata, year)

# Head and tail of data
head(ourdata)
tail(ourdata)

# Graph
library(ggplot2)
ggplot(ourdata) + 
  geom_line(aes(x=year,y=LE_fem,color='Female')) +
  geom_line(aes(x=year,y=LE_male,color='Male')) + 
  theme_light() + 
  labs(title="Life expectancy of females and males in China",
       subtitle="World Bank: World Development Indicators",
       x = "Year",
       y = "Life expectancy [years]",
       color = "Gender"
       ) 
```
`dplyr` also provides an equivalent way called the "pipe" (`%>%`)

```{r}
remove(ourdata)
ourdata <- wdi_raw %>% 
  # filter: only China data
    filter(iso2c=="CN") %>%
  # rename lifeexpectancy variables
    rename(LE_fem=SP.DYN.LE00.FE.IN) %>%
    rename(LE_male=SP.DYN.LE00.MA.IN) %>%
  # select relevant variables # help(select, package='dplyr')
    select(year, LE_fem, LE_male) %>%
  # order by year (increasing)
    arrange(year)

# Head and tail of data
head(ourdata)
tail(ourdata)

# Graph
library(ggplot2)
ggplot(ourdata) + 
  geom_line(aes(x=year,y=LE_fem,color='Female')) +
  geom_line(aes(x=year,y=LE_male,color='Male')) + 
  theme_light() + 
  labs(title="Life expectancy of females and males in China",
       subtitle="World Bank: World Development Indicators",
       x = "Year",
       y = "Life expectancy [years]",
       color = "Gender"
       ) 
```

Frequently, we merge two data sets on one or more common variable(s). This is called **merge** in Stata and **join** in R and Python. Think about how you would compute the B/M ratio for A-share stocks. 

```{r eval=FALSE}
le_data <- wdi_raw %>% rename(LE_fem=SP.DYN.LE00.FE.IN) %>% rename(LE_male=SP.DYN.LE00.MA.IN)
ctryinfo <- as.data.frame(WDI_data$country, stringsAsFactors = FALSE) %>% select(country, income)
save(le_data, ctryinfo, file='wdi_join.Rdata')
```

```{r}
load('wdi_join.Rdata')
tail(le_data)
tail(ctryinfo)
```
Join data set `le_data` and data set `ctryinfo` on variable `country`

```{r}
alldata <- left_join(le_data, ctryinfo, by = 'country')
tail(alldata)
```

Generate a new data set of life expectancies of income groups over time. This is called `collapse` in Stata and `summarize` in R. 

```{r}
# Summarize by country and year
avgdata <- alldata %>%     
  filter(income != "Aggregates") %>%           # remove rows for aggregates
  filter(income != "Not classified") %>%       # remove unclassified ctries
  filter(!is.na(income)) %>%                     # remove missing ctries
  group_by(income, year) %>%                   # group by income classification
  summarize(LE_fem_avg  = mean(LE_fem , na.rm=TRUE),
            LE_male_avg = mean(LE_male, na.rm=TRUE)) %>% # average by group 
  ungroup()                                    # remove grouping
tail(avgdata)
```
```{r}
ggplot(avgdata, aes(year, LE_fem_avg, color=income)) +
  scale_color_grey() +                              # gray scale
  geom_line() 
```
The income groups are not ordered correctly in the above graph. To perfect it, we need to understand `factor` in R.

```{r}
unique(alldata$income)
unique(avgdata$income)
```

```{r echo=FALSE}
# install.packages('epiDisplay')
library(epiDisplay, include.only='tab1')
```

```{r}
tab1(alldata$income, graph=FALSE)
tab1(avgdata$income, graph=FALSE)
```

R can `factor` the income groups into a more meaningful ordering. 

```{r}
(不考)
# Order the levels meaningfully
avgdata$income_factorized <- factor( avgdata$income, 
                          levels = c("High income",
                                     "Upper middle income",
                                     "Lower middle income",
                                     "Low income") )
# Plot
ggplot(avgdata, aes(year, LE_fem_avg, color=income_factorized)) +
  geom_line(size=1) +                               # thicker lines 
  scale_color_grey() +                              # gray scale
  scale_x_continuous(breaks=seq(1960,2015,10)) +    # adjust x axis breaks  
  theme_light() +                                   # light theme (white background,...)
  labs(title="Life expectancy of women",            
       subtitle="Average by country classification",
       x="Year", y="Life expectancy [Years]",
       color="Income level",
       caption="Source: World Bank, WDI")
```

dplyr relocate